---
title: "hw02: Exploring Gapminder with dplyr"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gapminder)
library(DT)
library(knitr)
library(ggplot2)
```

## Exercise 1: Basic `dplyr`
### 1.1
Subset gapminder into three countries of choice in the 1970s using `filter()`.

```{r}
knitr::kable(gapminder %>% 
    filter(year < 1980 & year > 1970,
         country == 'Morocco' | country == 'Iran' | country == 'Japan'))
  
```



### 1.2
Select `country` and `gdpPercap` using piping from filtered dataset.

```{r}
knitr::kable(gapminder %>% 
    filter(year < 1980 & year > 1970,
         country == 'Morocco' | country == 'Iran' | country == 'Japan') %>% 
    select(country, gdpPercap))
```

### 1.3
Create a new variable in gapminder showing differences in life expectancy in a country. Select any countries that have experienced a drop in life expectancy.

```{r}
DT::datatable(gapminder %>% 
                group_by(country) %>% 
                mutate(lifeExpDiff = lifeExp-lag(lifeExp)) %>% 
                filter(lifeExpDiff < 0))
        
```
### 1.4
Organize gapminder to contain the max GDP per capita experienced by each country.

```{r}
DT::datatable(gapminder %>% 
               group_by(country) %>% 
               filter(gdpPercap == max(gdpPercap)))

```

### 1.5
Create a scatter plot of Canada's life expectancy vs GDP per capita.

```{r}
gapminder %>% 
  filter(country == 'Canada') %>% 
  ggplot(aes(x = lifeExp, y = gdpPercap))+
    geom_point() +
    labs(
      x = "Life Expectancy",
      y = "GDP per capita",
      title = "Canada's Life Expectancy vs GDP Per Capita"
    )
```

## Exercise 2: Exploring individual variables with `dplyr`

### Categorical Variable: __Continent__
#### Possible values:
```{r}
# Possible continent values and number of coutries in each continent (based on data from 2007).
knitr::kable(contCount<-gapminder %>% 
              filter(year == 2007) %>% 
              count(continent))
```

#### Typical values, spread, distribution:

```{r}
#Distribution of countries in each continent:
contCount %>% 
  ggplot(aes(x = continent, y = n))+
    geom_bar(stat = "identity") +
    labs(
      x = "Continent",
      y = "Number of Countries",
      title = "Number of Countries in Each Continent",
      subtitle = "(based upon data from gapminder in 2007)"
    )


#Average GDP in each continent in 2007
DT::datatable(gapminder %>% 
                group_by(continent, year) %>% 
	              filter(year == 2007) %>% 
                mutate(meanGDP = mean(gdpPercap)) %>%
                ungroup() %>% 
                group_by(meanGDP) %>% 
                arrange(meanGDP) %>% 
                distinct(continent, meanGDP))
                
                
```

### Quantitative Variable: __Population__
#### Possible values:

Possible values of population across the years presented in gapminder:
```{r}
#absolute minimum, maximum and standard deviation:
knitr::kable(gapminder %>% 
              mutate(minimum = min(pop), maximum = max(pop), spread = sd(pop)) %>% 
              distinct(minimum, maximum, spread))
```


Population trends over time:

```{r}
#Global Population over the years
gapminder %>% 
  group_by(year) %>% 
  summarize(pop=sum(pop)/1) %>% 
  ggplot(aes(x = year, y = pop))+
    geom_point() +
    labs(
      x = "Year",
      y = "Population",
      title = "Global Population Over Time",
      subtitle = "(based upon data from gapminder from 1952-2007)"
    )
```

## Exercise 3: Exploring plots with `ggplot2`

### Scatterplot of 2 quantitative variables:

```{r}
gdp02 <- gapminder %>% 
  filter(year==2002) %>% 
  select(year, country, gdpPercap)

gdp77 <- gapminder %>% 
  filter(year==1977) %>% 
  select(year, country, gdpPercap)

gdpcomp <- tibble(country = (gdp77$country), g77 = (gdp77$gdpPercap), g02 = (gdp02$gdpPercap))

gdpcomp %>% 
  ggplot(aes(x = g77, y = g02))+
    geom_point(alpha = 0.4) +
    geom_abline(intercept = 0, slope = 1) +
    geom_text(aes(label = ifelse((g02>10000 | g77>10000),     as.character(country),''),hjust=0,vjust=0)) +
  labs(
      x = "GDP per capita in 1977",
      y = "GDP per capita in 2002",
      title = "GDP per capita in 1997 and 2002",
      subtitle = "(Countries above the black line experienced GDP per capita growth during this time period.)"
    )

```

### A different type of graph:

```{r}
#violin plot of gdp percap
qplot(continent, gdpPercap, data = gapminder, geom = "violin")
```

